---
title: "Distinguishing some relationships of asthma in the US from 2017-March 2020"
author: "Kelli Williams"
output: 
  html_document:
    theme: paper 
    highlight: tango
---
```{r eval = TRUE, message = FALSE}
library(haven)
library(nhanesA) 
library(plyr)
library(dplyr)
library(ggplot2)
library(cowplot)
#remotes::install_github("ddsjoberg/gtsummary")
library(gtsummary)
#devtools::install_github("sjPlot/devel")
library(sjPlot)
library(sjmisc)
library(sjlabelled)
```  
***
Use this template to complete your project throughout the course. Your Final Project presentation will be based on the contents of this document. Replace the title/name above and text below with your own, but keep the headers.

### Overview
Give a brief a description of your project and its goal(s), what data you are using to complete it, and what three faculty/staff in different fields you have spoken to about your project with a brief summary of what you learned from each person. Include a link to your final project GitHub repository.


### Introduction 
Describe the problem addressed, its significance, and some background to motivate the problem.

Explain why your problem is interdisciplinary, what fields can contribute to its understanding, and incorporate background related to what you learned from meeting with faculty/staff.


### Methods
The data sets utilized in this project were obtained from NHANES 2017-March 2020, which is a survey-based database that aims to assess the health and nutritional status of adults and children in the United States. Specifically, the questionnaires used were: demographics, health insurance, hospital utilization & access to care, and medical conditions. Several analyses will be done to address the three aims.

First, the raw data will be imported per NHANES questionnaire corresponding to 2017-March 2020. The raw data and codebook was downloaded from https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Questionnaire&Cycle=2017-2020.
```{r eval = TRUE, message = FALSE}
P_MCQ <- read_dta("NHANES original data files/P_MCQ.dta")
P_DEMO <- read_dta("NHANES original data files/P_DEMO.dta")
P_HIQ <- read_dta("NHANES original data files/P_HIQ.dta")
P_HUQ <- read_dta("NHANES original data files/P_HUQ.dta")
```
Then, for each of the four data sets, only relevant variables will be redefined in the cleaning of the raw data by using the NHANES code book. 
```{r eval = TRUE, message = FALSE}
P_MCQ2 <- P_MCQ %>%
  dplyr::mutate(mcq010 = factor(mcq010, levels =  c(1, 2), 
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq035 = factor(mcq035, levels = c(1, 2), 
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq040 = factor(mcq040, levels = c(1, 2), 
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq050 = factor(mcq050, levels = c(1, 2), 
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq160b = factor(mcq160b, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq160c = factor(mcq160c, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq160d = factor(mcq160d, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq160e = factor(mcq160e, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq160f = factor(mcq160f, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq160m = factor(mcq160m, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq160m = factor(mcq160p, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq160l = factor(mcq160l, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq500 = factor(mcq500, levels = c(1, 2), 
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq220 = factor(mcq220, levels = c(1, 2), 
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq371a = factor(mcq371a, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq371b = factor(mcq371b, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq371c = factor(mcq371c, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::mutate(mcq371d = factor(mcq371d, levels = c(1, 2), 
                                 labels =c("Yes", "No"))) %>%
  dplyr::rename(id = seqn, everasthma = mcq010, asthma = mcq035, 
                asthmaattack = mcq040, erasthma = mcq050, 
                chf =  mcq160b, chd = mcq160c, angina = mcq160d, 
                mi = mcq160e,stroke = mcq160f, everthyroid = mcq160m, 
                thyroid = mcq170m, lungdx = mcq160p, everliver = mcq160l, liver = mcq170l, 
                cancer = mcq220, losingweight = mcq371a, exercise = mcq371b, 
                reducesalt = mcq371c, reducefat = mcq371d) %>%
  dplyr::select(id, everasthma, asthma, asthmaattack, erasthma, 
                chf, chd, angina, mi, stroke, everthyroid, thyroid, lungdx, 
                everliver, liver, cancer, losingweight, exercise, reducesalt,
                reducefat)

# Create categorical variable from numeric variable indfmpir and label as number
# for later cleaning
P_DEMO$category[P_DEMO$indfmpir < 1] <- "Below the poverty level (<1)" <- 1
P_DEMO$category[P_DEMO$indfmpir == 1] <- "Equal to the poverty level (1)" <- 2
P_DEMO$category[P_DEMO$indfmpir > 1 & P_DEMO$indfmpir < 5] <-
                                  "Greater than the poverty level (>1)" <- 3

P_DEMO2 <- P_DEMO %>%
  dplyr::mutate(riagendr = factor(riagendr, 
                                  levels =  c(1, 2), labels =c("Male", "Female"))) %>%
  dplyr::mutate(ridreth3 = factor(ridreth3, 
                                  levels = c(1, 2, 3, 4, 5, 7), 
                                  labels = c("Mexican American", "Other Hispanic",
                                             "Non-Hispanic White", "Non-Hispanic Black",
                                             "Non-Hispanic Asian", 
                                             "Other Race- Including Multiracial"))) %>%
  dplyr::mutate(dmdeduc2 = factor(dmdeduc2, 
                                  levels = c(1, 2, 3, 4, 5), 
                                  labels = c("< 9th grade", "9th-11th grade",
                                             "High school graduate/GED", 
                                             "Some college or AA degree", 
                                             "College graduate or above"))) %>%
  dplyr::mutate(dmdmartz = factor(dmdmartz, 
                                  levels = c(1, 2, 3), 
                                  labels =c("Married/Living with Partner",
                                            "Widowed/Divorced/Separated", 
                                            "Single"))) %>%
  dplyr::mutate(ridageyr = dplyr::case_when(ridageyr <= 17 ~ "0-17",
                                            ridageyr > 17 ~ "18-80"),
                                            ridageyr = factor(ridageyr,
                                            levels = c("0-17","18-80"))) %>%
  dplyr::mutate(category = factor(category, 
                                  levels = c(1, 2, 3), 
                                  labels =c("Below the poverty level (<1)",
                                            "Equal to the poverty level (1)", 
                                            "Greater than the poverty level (>1)"))) %>%
  dplyr::rename(id = seqn, gender = riagendr, age = ridageyr, 
           race = ridreth3, education = dmdeduc2,
           maritalstatus = dmdmartz, ratio_poverty = category) %>%
  dplyr::select(id, gender, age, race, education, maritalstatus, ratio_poverty)

P_HIQ2 <- P_HIQ %>%
  dplyr::mutate(hiq011 = factor(hiq011, levels =  c(1, 2),
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(hiq032a = factor(hiq032a, levels = c(1),
                                 labels = c("Yes"))) %>%
  dplyr::mutate(hiq032b = factor(hiq032b, levels = c(2),
                                 labels = c("Yes"))) %>%
  dplyr::mutate(hiq032c = factor(hiq032c, levels = c(3),
                                 labels =c("Yes"))) %>%
  dplyr::mutate(hiq032d = factor(hiq032d, levels = c(4),
                                 labels =c("Yes"))) %>%
  dplyr::mutate(hiq032e = factor(hiq032e, levels = c(5),
                                 labels =c("Yes"))) %>%
  dplyr::mutate(hiq032h = factor(hiq032h, levels = c(8),
                                 labels =c("Yes"))) %>%
  dplyr::mutate(hiq032i = factor(hiq032i, levels = c(9),
                                 labels =c("Yes"))) %>%
  dplyr::mutate(hiq260 = factor(hiq260, levels = c(1, 2),
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(hiq270 = factor(hiq270, levels = c(1, 2),
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(hiq210 = factor(hiq210, levels = c(1, 2),
                                labels =c("Yes", "No"))) %>%
  dplyr::rename(id = seqn, insurancebinary = hiq011, privateinsurance = hiq032a, 
           medicare = hiq032b, medigap = hiq032c, medicaid = hiq032d, 
           chip = hiq032e, statesponsored = hiq032h, othergovernment = hiq032i, 
           havemedicare = hiq260, prescription_insurance = hiq270,
           noinsuranceyear = hiq210) %>%
  dplyr::select(id, insurancebinary, privateinsurance, medicare,
                medigap, medicaid, chip, statesponsored, othergovernment,
                havemedicare, prescription_insurance, noinsuranceyear)

P_HUQ2 <- P_HUQ %>%
  dplyr::mutate(hud062 = factor(hud062, levels =  c(1, 2, 3, 4),
                                labels =c("anytime less than 12 months ago",
                                          "1 year but less than 2 years ago",
                                          "within 3 years but less than 5 years",
                                          "within 5 years but less than 10 years"))) %>%
  dplyr::mutate(huq010 = factor(huq010, levels = c(1, 2, 3, 4, 5), 
                                labels = c("Excellent", "Very good", "Good",
                                           "Fair", "Poor"))) %>%
  dplyr::mutate(huq030 = factor(huq030, levels = c(1, 2, 3),
                                labels = c("Yes", "There is no place",
                                           "There is more than one place"))) %>%
  dplyr::mutate(huq051 = factor(huq051, levels = c(0, 1, 2, 3, 4, 5, 6, 7, 8),
                                labels =c("None", "1", "2 to 3", "4 to 5",
                                          "6 to 7", "8 to 9", "10 to 12",
                                          "13 to 15", "16 or more"))) %>%
  dplyr::mutate(huq071 = factor(huq071, levels = c(1, 2),
                                labels =c("Yes", "No"))) %>%
  dplyr::mutate(huq090 = factor(huq090, levels = c(1, 2),
                                labels =c("Yes", "No"))) %>%
  dplyr::rename(id = seqn, combinationdoctor = hud062, healthcondition = huq010, 
           routinehealthcare = huq030, numberhealthcare = huq051, 
           overnighthospital = huq071, mentalhealthvisit = huq090) %>%
  dplyr::select(id, combinationdoctor, healthcondition, routinehealthcare,
                numberhealthcare, overnighthospital, mentalhealthvisit)
```
The four data sets are then merged based on participant identifier into one data set.
```{r eval = TRUE, message = FALSE}
nhanes2017 <- Reduce(function(x, y) merge (x = x, y = y, all = TRUE, by = "id"), 
                     list(P_MCQ2, P_DEMO2, P_HIQ2, P_HUQ2))
```
Summary views of the combined data  set were reviewed, including Table 1.
```{r}
str(nhanes2017)
#How many NA's for each variable?
sapply(nhanes2017, function(x) sum(is.na(x)))

nhanes2017_dem <- nhanes2017 %>%
  select(asthma, everasthma, asthmaattack, erasthma, gender, age, race, education,
         maritalstatus, ratio_poverty, insurancebinary, privateinsurance, medicare,
         medigap, medicaid, chip, statesponsored, othergovernment, havemedicare,
         prescription_insurance, noinsuranceyear)
nhanes2017_dem$asthma <- forcats::fct_explicit_na(nhanes2017_dem$asthma)
table1 <- tbl_summary(nhanes2017_dem, by = asthma, missing = "no") %>%
  add_n() %>%
  add_overall() %>%
  add_stat_label() %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Table 1. Decriptive Characteristics by Asthma Status**") %>%
  bold_labels
```
Asthma prevalence was defined as ever being told you have asthma and having an asthma attack in the past year. A new data frame was created for asthma prevalence, as well as to use moving forward.
```{r eval = TRUE, message = FALSE}
nhanes2017_prev <- nhanes2017 %>%
  dplyr::select(everasthma, asthmaattack, gender, age, race, education, maritalstatus,
         ratio_poverty, insurancebinary, privateinsurance, medicare, medigap, medicaid,
         chip, statesponsored, othergovernment, havemedicare,
         prescription_insurance, noinsuranceyear, erasthma, combinationdoctor,
         healthcondition, routinehealthcare,numberhealthcare, overnighthospital,
         mentalhealthvisit) %>%
  dplyr::filter(everasthma == "Yes") %>%
  dplyr::filter(asthmaattack == "Yes")
nhanes2017_prev = data.frame(nhanes2017_prev)
```
To get an overall summary of those with asthma, Table 2 was created, as well as ####GRAPHS###
```{r eval = TRUE, message = FALSE}
table2 <- tbl_summary(nhanes2017_prev, by = everasthma, missing = "no") %>%
  add_n() %>%
  add_overall() %>%
  add_stat_label() %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Table 1. Decriptive Characteristics by Those with Asthma**") %>%
  bold_labels

#Asthma Prevalence table & graphs: social factors
# p1a <- ggplot2::ggplot(data = nhanes2017_prev, aes(x = race,
#                               fill = gender)) +
#                               geom_bar(position = "dodge") + ylab("Percent")
# p2a <- ggplot2::ggplot(data = nhanes2017_prev, aes(x = race, y = ratio_poverty)) +
#                                geom_boxplot()
# p2ab <- ggplot2::ggplot(data = nhanes2017_prev, aes(x = race, y = ratio_poverty)) +
#                                geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))
# p3a <- ggplot2::ggplot(data = nhanes2017_prev, aes(x = ratio_poverty)) + 
#                               geom_histogram(binwidth = 1)
# p4a <- ggplot2::ggplot(data = nhanes2017_prev, aes(x = insurancebinary,
#                               fill = factor(noinsuranceyear))) +
#                               geom_bar(position = "fill") + ylab("Percent")
#plot_grid(p1a, p2a, p2ab, p3a, p4a, labels = "AUTO")
```
A multivariate linear regression model was created to look at emergency room visits in those with asthma.
```{r eval = TRUE, message = FALSE}
#don't work= maritalstatus + education
#fit_glm1 <- glm(data = nhanes2017_prev, erasthma ~ race +  gender + ratio_poverty
                + age + maritalstatus + education + insurancebinary,
                family = binomial())
#works
fit_glm1 <- glm(data = nhanes2017_prev, erasthma ~ race + gender + ratio_poverty + age + insurancebinary 
                , 
                family = binomial())
summary(fit_glm1)
anova(fit_glm1)
tab_model(fit_glm1)
#need to fix labeling issue
# table_glm1 <- tab_model(
#   fit_glm1, 
#   pred.labels = c("Intercept", "Race", "Gender (Female)", "Poverty Score", "Age", "Marital Status", "Education", "Have Insurance (No)"),
#   dv.labels = c("multivariate linear regression model"),
#   string.pred = "Coeffcient",
#   string.ci = "Conf. Int (95%)",
#   string.p = "P-Value"
# )
# 
# summary(glm(data = nhanes2017_prev, erasthma ~ race + gender,
#               family = binomial()))
```
A multivariate linear regression model was created to look at healthcare provider and emergency room visits in those with asthma.
```{r eval = TRUE, message = FALSE}
fit_glm2 <- glm(data = nhanes2017_prev, erasthma ~ numberhealthcare +
                  mentalhealthvisit + overnighthospital + race + gender +
                  ratio_poverty + age + maritalstatus + education + insurancebinary,
                family = binomial())
summary(fit_glm2)
anova(fit_glm2)
tab_model(fit_glm2)
#both models side by side
tab_model(fit_glm1, fit_glm2)
```
### Results
Describe your results and include relevant tables, plots, and code/comments used to obtain them. End with a brief conclusion of your findings related to the question you set out to address. You can include references if you'd like, but this is not required.

In the first aim, we looked at asthma prevalence from 2017-March 2020 in the US. 
```{r}

```
In the second aim, we looked at emergency room visits in those with asthma from 2017-March 2020 in the US.
```{r}

```
In the third aim, we looked at healthcare provider visits and emergency room visits in those with asthma from 2017-March 2020 in the US.
```{r}

```



#############################DELETE BELOW##################################################
indfmpir < 1 ~ "Below the poverty level (<1)",
                                            indfmpir = 1 ~ "Equal to the poverty level (1)",
                                            indfmpir > 1 ~ "Greater than the poverty level (>1)"),
                                            indfmpir = factor(indfmpir,
                                            levels = c("<1","1", ">1")))

GLM tables (https://cran.r-project.org/web/packages/sjPlot/vignettes/tab_model_estimates.html)
#View of some variables
head(nhanes2017)
count_unique <- rapply(nhanes2017,function(x) length(unique(x)))
#There are 15,560 subjects in dataset
table(nhanes2017$gender) #males = 7721 & females = 7839
#table(nhanes2017$huq062) # "Combination of HUQ061 and HUQ062"
table(nhanes2017$everasthma) #number of "ever been told you have asthma" = 2322
table(nhanes2017$asthmaattack) #number of "Had asthma attack in past year" = 636
table(nhanes2017$asthma) #number of "still have asthma" = 1423
table(nhanes2017$erasthma) #number of "ER visit for asthma/past yr" = 295
summary(nhanes2017$age) #ages

#Table 1: Descriptive table: https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html; https://cran.r-project.org/web/packages/Gmisc/vignettes/Descriptives.html; https://www.danieldsjoberg.com/gtsummary/


#Brief overview graphs
  ####work on
p1 <- ggplot2::ggplot(data = nhanes2017, aes(x = gender,
                              fill = factor(asthma))) +
                              geom_bar(position = "fill") + ylab("Percent")
p2 <- ggplot2::ggplot(data = nhanes2017, aes(x = race,
                              fill = factor(asthma))) +
                              geom_bar(position = "fill") + ylab("Percent")
p3 <- ggplot2::ggplot(data = nhanes2017, aes(x = ratio_poverty,
                              y = factor(asthma)))geom_point(position = "fill")
p4 <- ggplot2::ggplot(data = nhanes2017, aes(x = insurancebinary,
                              fill = factor(asthma))) +
                              geom_bar(position = "fill") + ylab("Percent")
plot_grid(p1, p2, p3, p4, labels = "AUTO")


#GLM Tests
#gender <- drop_na(nhanes2017$gender)
# ggplot(data = nhanes2017, aes(x =erasthma, fill = gender)) +
#   geom_bar(position = "fill") 
# glm_a <- summary(glm(data = nhanes2017, erasthma ~ gender,
#                      family = binomial(link = "logit")))
# glm_b <- summary(glm(data = nhanes2017, erasthma ~ race,
#                      family = binomial(link = "logit")))
# chisq.test(table(nhanes2017$erasthma, nhanes2017$race))
# exp(coef(glm(erasthma ~ race, data = nhanes2017, family = binomial()))) 

# nhanes2017$poverty = as.numeric(nhanes2017$poverty)
# nhanes2017$age = as.numeric(nhanes2017$age)
# nhanes2017$erasthma = as.numeric(nhanes2017$erasthma)
# nhanes2017$race = as.numeric(nhanes2017$race)
# nhanes2017$maritalstatus = as.numeric(nhanes2017$maritalstatus)
# nhanes2017$education = as.numeric(nhanes2017$education)
# nhanes2017$insurancebinary = as.numeric(nhanes2017$insurancebinary)
# nhanes2017$prescription_insurance = as.numeric(nhanes2017$prescription_insurance)
#prescription_insurance makes it not run
#fit_glm <- glm(data = nhanes2017_prev, erasthma ~ race + gender + poverty + age +
               #maritalstatus + education + insurancebinary + prescription_insurance, family = binomial())


#works, need to problem solve above solution for the variables that won't run as is
#class(nhanes2017_prev$combinationdoctor)
#overnighthospital, privateinsurance, havemedicare noinsuranceyear,
# nhanes2017$numberhealthcare = as.numeric(nhanes2017$numberhealthcare)
# nhanes2017$combinationdoctor = as.numeric(nhanes2017$combinationdoctor)
# nhanes2017$mentalhealthvisit = as.numeric(nhanes2017$mentalhealthvisit)
# nhanes2017$overnighthospital = as.numeric(nhanes2017$overnighthospital)
# nhanes2017$privateinsurance = as.numeric(nhanes2017$privateinsurance)
# nhanes2017$havemedicare = as.numeric(nhanes2017$havemedicare)
# nhanes2017$noinsuranceyear = as.numeric(nhanes2017$noinsuranceyear)
